Normalization by Evaluation for Sized Dependent Types

### 2017-02-23 nbe/icfp17.tex
[ ICFP 17 ] Some refinement of the Fridlender/Pagano theorems.

\nextcase $A' \evalsto \piT A B$ and $\resurrect\Gamma \der T' = \piT U T$.
\begin{enumerate}
\item To demonstrate $\LR \Gamma t {T'} {\up {A'} n} {A'}$ we assume
$\xi : \Gamma' \leq \Gamma$ and $\LR{\Gamma'}{u}{U\xi} a A$ and show
$\LR{\Gamma'} {(t\xi\,u)} {T(\xi,u)} {\up{B[a]}(n\,\down A a)} {B[a]}$.
This follows by induction hypothesis (\ref{it:into}) for $B[a] \in \SET[\ell]$ if
we can prove that for all $\xi' : \Gamma'' \leq \Gamma'$ there is a neutral normal form $m$
such that $\RNe{\Gamma''} {(n\,\down A a)} m$ and $\Gamma'' \der (t\xi\,u)\xi' = m : T(\xi,u)\xi'$.
Note that the last goal can be rewritten to $\Gamma'' \der t\xi\xi'\,(u\xi') = m : T(\xi\xi',u\xi')$.
We know that $\RNe{\Gamma''} n {m_0}$ with $\Gamma'' \der t\xi\xi' = m_0 : T'\xi\xi'$.
Thus, it is sufficient to show that there is some normal form $v$ such that
$\R{\Gamma''} {\down A a} v$ and $\Gamma'' \der u\xi' = v : U\xi\xi'$.
However, this follows by induction hypothesis (\ref{it:out}) on
assumption $\LR{\Gamma'}{u}{U\xi} a A$ with $\xi'$.
\end{enumerate}
\end{caselist}
\end{proof}
\begin{corollary}[One-to-one]
  If $\LRT \Gamma T A \ell$ and $\LRT \Gamma {T'} A {\ell'}$ then $\Gamma \der T = T'$.
\end{corollary}

@@

\begin{lemma}[\citet{fridlenderPagano:tlca13}]
   If $\LRT \Gamma T A \ell$ and $\LRT \Gamma {T'} {A'} {\ell'}$ and $A \leq A' \in \TYPE$ then $\Gamma \der T \leq T'$.
\end{lemma}

\begin{lemma}[\citet{fridlenderPagano:tlca13}]
   If $\LRT \Gamma T A \ell$ and $\LRT \Gamma {T'} {A'} {\ell'}$ and $A \leq A' \in \TYPE$ then
   $\LR \Gamma t T f A$ implies $\LR \Gamma t {T'} f {A'}$.
\end{lemma}

>>>>

\nextcase $A' \evalsto \piT A B$ and $\resurrect\Gamma \der T' = \piT U T$.
\begin{enumerate}
\item To demonstrate $\LR \Gamma t {T'} {\up {A'} n} {A'}$ we assume
$\xi : \Gamma' \leq \Gamma$ and $\LR{\Gamma'}{u}{U\xi} a A$ and show
$\LR{\Gamma'} {(t\xi\,u)} {T(\xi,u)} {\up{B[a]}(n\,\down A a)} {B[a]}$.
This follows by induction hypothesis (\ref{it:into}) for $B[a] \in \SET[\ell]$ if
we can prove that for all $\xi' : \Gamma'' \leq \Gamma'$ there is a neutral normal form $m$
such that $\RNe{\Gamma''} {(n\,\down A a)} m$ and $\Gamma'' \der (t\xi\,u)\xi' = m : T(\xi,u)\xi'$.
Note that the last goal can be rewritten to $\Gamma'' \der t\xi\xi'\,(u\xi') = m : T(\xi\xi',u\xi')$.
We know that $\RNe{\Gamma''} n {m_0}$ with $\Gamma'' \der t\xi\xi' = m_0 : T'\xi\xi'$.
Thus, it is sufficient to show that there is some normal form $v$ such that
$\R{\Gamma''} {\down A a} v$ and $\Gamma'' \der u\xi' = v : U\xi\xi'$.
However, this follows by induction hypothesis (\ref{it:out}) on
assumption $\LR{\Gamma'}{u}{U\xi} a A$ with $\xi'$.
\end{enumerate}
\end{caselist}
\end{proof}
\begin{corollary}[One-to-one]
\bla
% \begin{enumerate}
% \item
% If $\LRSize \Gamma a \alpha$ and $\LRSize \Gamma {a'} \alpha$ then $a = a'$.
% \item
If $\LRT \Gamma T A \ell$ and $\LRT \Gamma {T'} A {\ell'}$ then $\Gamma \der T = T'$.
%\end{enumerate}
\end{corollary}

@@

\begin{lemma}[Semantic subtyping implies judgmental subtyping \cite{fridlenderPagano:tlca13}]
\bla
\begin{enumerate}
\item
   If $\LRSize \Gamma a \alpha$ and $\LRSize \Gamma {b} \beta$ and $\alpha \leq \beta$ then $a \leq b$.
\item
   If $\LRT \Gamma T A \ell$ and $\LRT \Gamma {T'} {A'} {\ell'}$ and $A \leq A' \in \TYPE$ then $\Gamma \der T \leq T'$.
\end{enumerate}
\end{lemma}

\begin{lemma}[Subsumption for the logical relation \cite{fridlenderPagano:tlca13}]
   If $\LRT \Gamma T A \ell$ and $\LRT \Gamma {T'} {A'} {\ell'}$ and $A \leq A' \in \TYPE$ then
   $\LR \Gamma t T f A$ implies $\LR \Gamma t {T'} f {A'}$.
\end{lemma}
